---
title: 'Evaluating Client Presentation Effectiveness in Retail Banking'
author: "Your Name"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(caret)
library(class)
library(e1071)
library(rpart)
library(rpart.plot)
library(factoextra)
```

## 1. Load and Check Data

```{r}
data <- read_excel("Bank_dataset.xlsx", sheet = 1)

print(head(data))

print(colnames(data))

print(dim(data))
```

## 2. Data Preparation and Conversion

```{r}
data$Presentation <- as.factor(data$Presentation)
data$Gender <- as.factor(data$Gender)

names(data) <- str_trim(names(data))

print(head(data))
```

## 3. Exploratory Data Analysis (EDA)

```{r}
ggplot(data, aes(x = Presentation, y = Profit, fill = Presentation)) +
  geom_boxplot() +
  labs(title = "Profit by Presentation", y = "Profit")

ggplot(data, aes(x = Gender, y = Profit, fill = Presentation)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Profit by Gender and Presentation", y = "Profit")
```

## 4. Regression Analysis (Supervised Learning)

```{r}
model_lm <- lm(Profit ~ Presentation + Age + Gender + `Monthly Income`, data = data)
summary(model_lm)
```

## 5. K-Nearest Neighbors (KNN Classification)

```{r}
df_knn <- data %>%
  mutate(Gender = as.numeric(Gender)) %>%
  select(Presentation, Age, Gender, `Monthly Income`, Profit) %>%
  drop_na()

df_knn[, 2:5] <- scale(df_knn[, 2:5])

set.seed(123)
train_idx <- sample(nrow(df_knn), 0.7 * nrow(df_knn))
train <- df_knn[train_idx, ]
test <- df_knn[-train_idx, ]

knn_pred <- knn(train[, 2:5], test[, 2:5], train$Presentation, k = 5)

confusionMatrix(knn_pred, test$Presentation)
```

## 6. Decision Tree

```{r}
tree_model <- rpart(Presentation ~ Age + Gender + `Monthly Income` + Profit, data = data, method = "class")
rpart.plot(tree_model, type = 2, extra = 104, fallen.leaves = TRUE)
```

## 7. Cluster Analysis (Unsupervised Learning)

```{r}
df_cluster <- data %>%
  select(Age, `Monthly Income`, Profit) %>%
  drop_na()

df_scaled <- scale(df_cluster)

set.seed(123)
kmeans_model <- kmeans(df_scaled, centers = 3)
data$Cluster <- factor(kmeans_model$cluster)

fviz_cluster(kmeans_model, data = df_scaled)
```

## 8. Conclusion

This case study used both supervised and unsupervised learning techniques to investigate why clients who received a product presentation had lower average profitability.

**Findings:**
- Regression shows potential confounding variables such as age, gender, or income.
- KNN and decision trees attempt classification based on client profile.
- Clustering may reveal hidden client segments with distinct behaviors.
