---
title: 'Analyzing Retail Banking Metrics Based on Agent and Payment Characteristics'
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(caret)
library(class)
library(e1071)
library(rpart)
library(rpart.plot)
library(factoextra)
```

## 1. Load and Check Data

```{r}
# Load the dataset
data <- read_excel("Business.xlsx", sheet = 2)

# Basic checks
print(head(data))
print(colnames(data))
print(dim(data))
```

## 2. Data Preparation and Conversion

```{r}
# Clean column names
names(data) <- str_trim(names(data))

# Convert relevant columns to factors
data$Agent_Cat <- as.factor(data$Agent_Cat)
data$Agent_Channel <- as.factor(data$Agent_Channel)
data$Mode_of_Pay <- as.factor(data$Mode_of_Pay)
data$ECS_Status <- as.factor(data$ECS_Status)
data$City_Tier <- as.factor(data$City_Tier)
data$Fund_Type <- as.factor(data$Fund_Type)

# View cleaned data
print(head(data))
```

## 3. Exploratory Data Analysis (EDA)

```{r}
# Premium Amount by Agent Category
ggplot(data, aes(x = Agent_Cat, y = Premium_Amt, fill = Agent_Cat)) +
  geom_boxplot() +
  labs(title = "Premium Amount by Agent Category", y = "Premium Amount", x = "Agent Category")

# Cibil Score by Mode of Payment
ggplot(data, aes(x = Mode_of_Pay, y = Cibil_Score, fill = Mode_of_Pay)) +
  geom_boxplot() +
  labs(title = "Cibil Score by Mode of Payment", y = "Cibil Score", x = "Mode of Payment")
```

## 4. Linear Regression (Supervised Learning)

```{r}
# Linear model to predict Cibil_Score
model_lm <- lm(Cibil_Score ~ Premium_Amt + Mean_NDPD + Agent_Cat + Mode_of_Pay + City_Tier, data = data)
summary(model_lm)
```

## 5. Logistic Regression (Supervised Learning)

```{r}
# Create binary ECS Status only if ECS_Status exists and is not NA
if (!is.null(data$ECS_Status) && any(!is.na(data$ECS_Status))) {

  # Make sure ECS_Status_Binary has 0 and 1
  data$ECS_Status_Binary <- ifelse(data$ECS_Status == "Active", 1, 0)
  
  # Drop rows with NA in important columns
  data_logit <- data %>%
    filter(!is.na(ECS_Status_Binary), !is.na(Premium_Amt), !is.na(Mean_NDPD), !is.na(Cibil_Score))
  
  # Check if there are both classes present
  if (length(unique(data_logit$ECS_Status_Binary)) > 1) {
    
    # Logistic regression model
    model_logit <- glm(ECS_Status_Binary ~ Premium_Amt + Mean_NDPD + Cibil_Score + Agent_Cat + Mode_of_Pay + City_Tier,
                       data = data_logit, family = binomial)
    
    # Summary
    summary(model_logit)
    
    # Predict probabilities
    pred_probs <- predict(model_logit, type = "response")
    
    # Create predicted labels
    pred_labels <- ifelse(pred_probs > 0.5, 1, 0)
    
    # Confusion Matrix only if both classes are present
    if (length(unique(pred_labels)) > 1 & length(unique(data_logit$ECS_Status_Binary)) > 1) {
      confusionMatrix(as.factor(pred_labels), as.factor(data_logit$ECS_Status_Binary))
    } else {
      print("Not enough classes in predicted or actual labels to compute confusion matrix.")
    }
    
  } else {
    print("Only one class in ECS_Status_Binary. Logistic Regression is not meaningful.")
  }
  
} else {
  print("ECS_Status column is missing or empty. Skipping Logistic Regression.")
}
```

## 6. K-Nearest Neighbors (KNN Classification)

```{r}
# Prepare data for KNN
df_knn <- data %>%
  mutate(Agent_Cat = as.numeric(Agent_Cat),
         Agent_Channel = as.numeric(Agent_Channel),
         Mode_of_Pay = as.numeric(Mode_of_Pay),
         ECS_Status = as.numeric(ECS_Status),
         City_Tier = as.numeric(City_Tier),
         Fund_Type = as.numeric(Fund_Type)) %>%
  select(Premium_Amt, Mean_NDPD, Cibil_Score, Agent_Cat, Mode_of_Pay) %>%
  drop_na()

# Scale data
df_knn <- scale(df_knn)

# Split into training and testing sets
set.seed(123)
train_idx <- sample(nrow(df_knn), 0.7 * nrow(df_knn))
train <- df_knn[train_idx, ]
test <- df_knn[-train_idx, ]

# Train labels
train_labels <- data$Agent_Cat[train_idx]
test_labels <- data$Agent_Cat[-train_idx]

# KNN model
knn_pred <- knn(train, test, train_labels, k = 5)

# Confusion matrix
confusionMatrix(knn_pred, test_labels)
```

## 7. Decision Tree

```{r}
# Build decision tree model
tree_model <- rpart(Agent_Cat ~ Premium_Amt + Mean_NDPD + Cibil_Score + Mode_of_Pay + City_Tier, 
                    data = data, method = "class")

# Plot decision tree
rpart.plot(tree_model, type = 2, extra = 104, fallen.leaves = TRUE)
```

## 8. Cluster Analysis (Unsupervised Learning)

```{r}
# Prepare data for clustering
df_cluster <- data %>%
  select(Premium_Amt, Mean_NDPD, Cibil_Score) %>%
  drop_na()

# Scale the data
df_scaled <- scale(df_cluster)

# KMeans clustering
set.seed(123)
kmeans_model <- kmeans(df_scaled, centers = 3)

# Assign cluster labels
data$Cluster <- factor(kmeans_model$cluster)

# Plot clusters
fviz_cluster(kmeans_model, data = df_scaled)
```

## 9. Conclusion

This case study applied both supervised and unsupervised learning methods:

- **Linear Regression** analyzed predictors of credit scores (Cibil Score).
- **Logistic Regression** modeled probability of active ECS_Status.
- **KNN Classification** predicted Agent Categories.
- **Decision Trees** visualized customer segmentation rules.
- **Clustering** revealed hidden groups based on payment and credit behavior.
